---
title: "Crime"
author: "Nachiket Hinge"
date: "3/10/2022"
output: pdf_document
---
# This is a 2018 Dataset

# Importing and Manipulating Dataset
```{r echo=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
georgia <- read_excel("Datasets/georgia.xls")
illinois <- read_excel("Datasets/illinois.xls")

#For Georgia
georgia <- georgia[c(5:258),]
colnames(georgia) <- c("City", "Population", "Violent.Crime", "Murder.And.NonNegligent.Manslaughter", "Rape", "Robbery", "Aggraveted.Assault", "Property.Crime", "Burglary", "Larceny.Theft", "Motor.Vehicle.Theft", "Arson")

#For Illinois
illinois <- illinois[c(5:525),]
colnames(illinois) <- c("City", "Population", "Violent.Crime", "Murder.And.NonNegligent.Manslaughter", "Rape", "Robbery", "Aggraveted.Assault", "Property.Crime", "Burglary", "Larceny.Theft", "Motor.Vehicle.Theft", "Arson")
```

# Checking Dimensions
```{r echo=FALSE}
print("Georgia: ")
dim(georgia)
print("Illinois: ")
dim(illinois)
```


Let's get the variables we are most interested in:

# Getting relevant Data for Georgia
```{r echo=FALSE}
georgia$Population <- as.numeric(georgia$Population)
georgia$Violent.Crime <- as.numeric(georgia$Violent.Crime)
georgia$Property.Crime <- as.numeric(georgia$Property.Crime)

new.georgia <- georgia %>% filter(Population > 5000) %>%  mutate(Crime.Rate = ((Violent.Crime + Property.Crime)/Population) * 100) %>% select(City, Crime.Rate) %>% arrange((Crime.Rate))
print(new.georgia)
```

# Analyzing Georgia
```{r}
which(new.georgia$City == "Alpharetta")
```


Alpharetta Ranks 17th in the list of least crimes out of 94 cities

```{r}
which(new.georgia$City == "Atlanta")
```

Atlanta does poorly in terms of crime.. It ranks 69th!!!
- Not Nice

```{r}
ga.cities <- c("Alpharetta", "Johns Creek", "Milton", "Duluth", "Decatur", "Savannah-Chatham Metropolitan", "Roswell", "Atlanta", "Kennesaw", "Sandy Springs")

filter(new.georgia, new.georgia$City %in% ga.cities)
```

Of the cities listed, Johns Creek is the safest followed by Milton
Atlanta is the least safest city in the list with a crime rate of 5.4%

```{r echo=FALSE}
ggplot(data = new.georgia) +
  geom_histogram(aes(x=Crime.Rate), color="white") +
  xlab("Crime Rate") +
  ggtitle("Typical Crime rate in Georgia") +
  xlim(0,15)
 # scale_x_continuous(breaks = scales::pretty_breaks(n = 15))
```

```{r}
new.georgia[which.max(new.georgia$Crime.Rate),]
```

East Point is the most dangerous city in Georgia

# Getting relevant Data for Illinois
```{r echo=FALSE}
illinois$Population <- as.numeric(illinois$Population)
illinois$Violent.Crime <- as.numeric(illinois$Violent.Crime)
illinois$Property.Crime <- as.numeric(illinois$Property.Crime)

new.illinois <- illinois %>% filter(Population > 5000) %>%  mutate(Crime.Rate = ((Violent.Crime + Property.Crime)/Population) * 100) %>% select(City, Crime.Rate) %>% arrange((Crime.Rate)) %>% na.omit
print(new.illinois)
```

```{r}
which(new.illinois$City == "Chicago")
```

Chicago can be considered as a dangerous city as it ranks 252th out of 277 cities in safety

# Analyzing Illinois
```{r echo=FALSE}
ggplot(data = new.illinois) +
  geom_histogram(aes(x=Crime.Rate), color="white") +
  xlab("Crime Rate") +
  ggtitle("Typical Crime rate in Illinois") +
  xlim(0, 15)
 
```

```{r}
new.illinois[which.max(new.illinois$Crime.Rate),]
```

North Riverside is the most dangerous city in Illinois

# Comparing Georgia and Illinois
```{r echo=FALSE}
print(paste("Average Crime Rate in Georgia", as.character(mean(new.georgia$Crime.Rate))))
print(paste("Average Crime Rate in Illinois", as.character(mean(new.illinois$Crime.Rate))))
```

```{r echo=FALSE}
atlanta <- new.georgia$Crime.Rate[new.georgia$City == "Atlanta"]
chicago <- new.illinois$Crime.Rate[new.illinois$City == "Chicago"]
print(paste("Crime rate in Atlanta:", as.character(atlanta)))
print(paste("Crime rate in Chicago:", as.character(chicago)))
```

```{r echo=FALSE}
ggplot() + 
  geom_density(aes(x = new.georgia$Crime.Rate), color = "white", fill="Red") + 
  geom_density(aes(x = new.illinois$Crime.Rate), color = "white", fill="Blue") +
  xlab("Crime Rate (Red = GA, Blue = IL)") +
  ylab("Density") +
  ggtitle("Crime Rate in Georgia vs Illinois")
```

```{r}
new.georgia$Crime.Rate[new.georgia$City == "Alpharetta"]
georgia$Population[georgia$City == "Alpharetta"]
```

```{r}
alpharetta.of.illinois <- new.illinois$City[new.illinois$Crime.Rate > 1.8 & new.illinois$Crime.Rate < 1.9]
alpharetta.of.illinois
```

```{r}
illinois %>% filter(City %in% alpharetta.of.illinois & Population > 50000 & Population < 80000)
```

Berwyn is the Alpharetta of Illinois


# Other Stuff
```{r}
il.cities <- c("Oak Brook", " South Barrington", "Lincolnwood", "Hoffman Estates", "Glendale Heights")

filter(new.illinois, new.illinois$City %in% il.cities)
```


